### 2020-04-30 TO_DO.Rmd

---
title: 'Goals and Tasks'
author: "ASP Boost Team"
date: "30 4 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I propose this file for technical discussion and goal setting and tracking. 
Goal setting could look like this (This is merely a sketch):

## Tasks to achieve till next Q&A session 20.05.04 
### (as set in the last meeting)

* ~~README~~ L. 
* ~~Rmd file for communication~~ S (this file)
* Push a boost function JLS.

Initials could indicate whom belongs which task.

Discussion could look like this (The html file is intended to visualize LaTeX):

## Estimator:

Residuals: $u = y = x\hat{\beta}$
Estimator *coming from the residuals*: $\hat{b} = (X'X)^{-1} X'u$

After inserting $u$ in $\hat{b}$ the base formula results:
$\hat{\beta}^{t}= \hat{\beta}^{t-1} + v(\hat{\beta} - \hat{\beta}^{t-1})$

I am unsure, if the first term inside parenthesis is just $\hat{\beta}$ or rather $\hat{\beta}^{something}$. Please check.