### 2020-05-12 README.Rmd
small change to readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

set.seed(1337)
```

# asp20boost

* Supervisor: Thomas Kneib
* Students: Johannes, Levin, Sebastián
* Topic: Parameter estimation and variable selection using the boosting ensemble learning algorithm

## Tasks

### until ~ 11 May

* simple boost function for scale (estimate var(yi))
* read Wickham "R-Packages"
* try implementing function-testing
* get used to roxygen2

## Literature
* Section **4.3** on "Boosting Linear Regression Models" in "Regression: Models, methods, and applications" by Ludwig Fahrmeir, Thomas Kneib, Stefan Lang, and Brian Marx
* Section **2.9.1** on "Regression models for location, scale, and shape" in ibd.

<!--
more datasets:
- gamlss.data::TODO
- MASS::Boston
- https://archive.ics.uci.edu/ml/datasets.php
-->

## Location Scale Regression Model

```{r data, fig.width = 11}
n <- 500
x <- runif(n)
y <- x + rnorm(n, sd = exp(-3 + 2 * x))
plot(x, y)
abline(0, 1, lwd = 2)
curve(x + 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
curve(x - 1.96 * exp(-3 + 2 * x), -0.1, 1.1, add = TRUE)
```

## Working with Location Scale Regression Boost Model
First, you need to install the `asp20model` &  `asp20boost` package from GitLab and load it:

```{r installation, eval = 3}
install.packages("devtools")
#This throws an error for asp20boost
#devtools::install_gitlab("asp20/asp20model", host = "gitlab.gwdg.de")
#devtools::install_gitlab("asp20/asp20boost", host = "gitlab.gwdg.de")

```

```{r}
library(asp20boost)
model <- LocationScaleRegressionBoost$new(y ~ x, ~ x)
gradient_boost(model, 
              stepsize = 0.001, maxit = 1000,
              abstol = 0.001,
              verbose = FALSE)
plot(x, y)
lines(x,model$fitted_location)
#Todo do create similar graphic
curve(x + 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)
curve(x - 1.96 * exp(x*model$gamma), -0.1, 1.1, add = TRUE)

```