### 2020-05-27 Reports/1_REPORT.Rmd
- Basic RMD structure for report. - JSS template for reference, if output desired (untitled). - Literature and packages and Bibtex files for citation. - First knit to pdf and to html

---
title: "1 Report ASP20 Boost"
subtitle: "First Report"
author:
- "Johannes Strauß"
- "Levin Wiebelt"
- "Sebastian Aristizabal"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: [nothing, nothingness]
abstract: |
  This is the abstract.

  It consists of two paragraphs.
documentclass: report
bibliography: ["citavi_27.05.bib", "packages.bib"]
biblio-style: apalike
geometry: "left=4cm, right=3cm, top=2.5cm, bottom=2.5cm"
lof: FALSE
lot: FALSE
output: 
    pdf_document:
      toc: true
      toc_depth: 2
      latex_engine: xelatex
      keep_tex: yes
      template: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# How to cite and configure the document's appearance

## Write

To cite:

* Parentheses: Implementation builds upon the asp20 model [@R-asp20model]. 
* Inline: It is implemented as R6 class @R-R6.
* Including pages: For the theoretical foundations of our implementation, we rely mostly on  [@Fahrmeir.2013, 219] and [@Hastie.2009, 358].

In the `report` folder there's two `.bib` files. 
These contain all reference we should need. 
To update the `package.bib` write the name of the package in the following chunk and *actively* run it (It should be installed in your terminal).

```{r pack tex, eval=FALSE, include=TRUE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'R6', 'knitr', 'rmarkdown', 'asp20model', 'gamboostLSS','mboost', 'tidyverse'  
), 'packages.bib')
```

To cite write a ``@` followed by `firstname.year` as shown in the `.bib` files for reference. 
See further [cite](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)

As a tip hit always enter after each period for better readability in R Studio.
Like this.

Two enters give me a new paragraph.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

## Appearance using the `YAML` Header:

We can regulate the appearance of the document from here. 
There's a lot flexibility but the learning curve is somehow steep.
I'll link useful resources to learn, but, in principle, the document is *ready for writing*.

Here the corresponding references:

*  [PDF output](https://bookdown.org/yihui/rmarkdown/pdf-document.html)
*  [Pandoc Manuals](https://pandoc.org/MANUAL.html#templates)
*  [TOC](https://bookdown.org/yihui/rmarkdown/pdf-document.html#table-of-contents-1)
*  [Cite with Pandoc](https://github.com/jgm/pandoc-citeproc/blob/master/man/pandoc-citeproc.1.md)

Here those inputs that are easy to configuration for you guys to get an idea:

* With `documentclass` has three base templates: `article`, `report` and `book`. 
* `geometry` let's me change the margins. 
* `lof` and `lot` set to true give me a list of figures and tables respectively.

# 0.Lead Questions to be answered (from Hannes' e-mail):

* Which open questions do you have about the statistical model and methodology?
*  What functionality are you planning to (or did you already) implement? What did you decide to leave out?
* What are the critical aspects of your implementation and your package? Stability? Performance? User-friendliness? How are you going to deal with these potential issues?
    What design designs did you make with respect to your implementation? How are you going to extend the R6 class? What additional functions are you going to provide?

# 1. Description of the problem

We have been working on an implementation for a boosting implementation with gradient descent for location and scale.

# 2. Description of progress

## Johannes is gzus.

# 3. Intended extensions to be implemented. 
# 4. Comments
# 5. References