### 2020-05-27 Reports/1_REPORT.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{report}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={1 Report ASP20 Boost},
  pdfauthor={Johannes Strauß; Levin Wiebelt; Sebastian Aristizabal},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=4cm, right=3cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{1 Report ASP20 Boost}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{First Report}
\author{Johannes Strauß \and Levin Wiebelt \and Sebastian Aristizabal}
\date{27 Mai 2020}

\begin{document}
\maketitle
\begin{abstract}
This is the abstract.

It consists of two paragraphs.
\end{abstract}

{
\setcounter{tocdepth}{1}
\tableofcontents
}
\hypertarget{how-to-cite-and-configure-the-documents-appearance}{%
\chapter{How to cite and configure the document's
appearance}\label{how-to-cite-and-configure-the-documents-appearance}}

\hypertarget{write}{%
\section{Write}\label{write}}

To cite:

\begin{itemize}
\tightlist
\item
  Parentheses: Implementation builds upon the asp20 model (Riebl 2020).
\item
  Inline: It is implemented as R6 class Chang (2019).
\item
  Including pages: For the theoretical foundations of our
  implementation, we rely mostly on (Fahrmeir 2013, 219) and (Hastie,
  Tibshirani, and Friedman 2009, 358).
\end{itemize}

In the \texttt{report} folder there's two \texttt{.bib} files. These
contain all reference we should need. To update the \texttt{package.bib}
write the name of the package in the following chunk and \emph{actively}
run it (It should be installed in your terminal).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# automatically create a bib database for R packages}
\NormalTok{knitr}\OperatorTok{::}\KeywordTok{write_bib}\NormalTok{(}\KeywordTok{c}\NormalTok{(}
  \KeywordTok{.packages}\NormalTok{(), }\StringTok{'R6'}\NormalTok{, }\StringTok{'knitr'}\NormalTok{, }\StringTok{'rmarkdown'}\NormalTok{, }\StringTok{'asp20model'}\NormalTok{, }\StringTok{'gamboostLSS'}\NormalTok{,}\StringTok{'mboost'}\NormalTok{, }\StringTok{'tidyverse'}  
\NormalTok{), }\StringTok{'packages.bib'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

To cite write a `\texttt{@} followed by \texttt{firstname.year} as shown
in the \texttt{.bib} files for reference. See further
\href{https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html}{cite}

As a tip hit always enter after each period for better readability in R
Studio. Like this.

Two enters give me a new paragraph.

\hypertarget{including-plots}{%
\section{Including Plots}\label{including-plots}}

You can also embed plots, for example:

\includegraphics{1_REPORT_files/figure-latex/pressure-1.pdf}

\hypertarget{appearance-using-the-yaml-header}{%
\section{\texorpdfstring{Appearance using the \texttt{YAML}
Header:}{Appearance using the YAML Header:}}\label{appearance-using-the-yaml-header}}

We can regulate the appearance of the document from here. There's a lot
flexibility but the learning curve is somehow steep. I'll link useful
resources to learn, but, in principle, the document is \emph{ready for
writing}.

Here the corresponding references:

\begin{itemize}
\tightlist
\item
  \href{https://bookdown.org/yihui/rmarkdown/pdf-document.html}{PDF
  output}
\item
  \href{https://pandoc.org/MANUAL.html\#templates}{Pandoc Manuals}
\item
  \href{https://bookdown.org/yihui/rmarkdown/pdf-document.html\#table-of-contents-1}{TOC}
\item
  \href{https://github.com/jgm/pandoc-citeproc/blob/master/man/pandoc-citeproc.1.md}{Cite
  with Pandoc}
\end{itemize}

Here those inputs that are easy to configuration for you guys to get an
idea:

\begin{itemize}
\tightlist
\item
  With \texttt{documentclass} has three base templates:
  \texttt{article}, \texttt{report} and \texttt{book}.
\item
  \texttt{geometry} let's me change the margins.
\item
  \texttt{lof} and \texttt{lot} set to true give me a list of figures
  and tables respectively.
\end{itemize}

\hypertarget{lead-questions-to-be-answered-from-hannes-e-mail}{%
\chapter{0.Lead Questions to be answered (from Hannes'
e-mail):}\label{lead-questions-to-be-answered-from-hannes-e-mail}}

\begin{itemize}
\tightlist
\item
  Which open questions do you have about the statistical model and
  methodology?
\item
  What functionality are you planning to (or did you already) implement?
  What did you decide to leave out?
\item
  What are the critical aspects of your implementation and your package?
  Stability? Performance? User-friendliness? How are you going to deal
  with these potential issues? What design designs did you make with
  respect to your implementation? How are you going to extend the R6
  class? What additional functions are you going to provide?
\end{itemize}

\hypertarget{description-of-the-problem}{%
\chapter{1. Description of the
problem}\label{description-of-the-problem}}

We have been working on an implementation for a boosting implementation
with gradient descent for location and scale.

\hypertarget{description-of-progress}{%
\chapter{2. Description of progress}\label{description-of-progress}}

\hypertarget{johannes-is-gzus.}{%
\section{Johannes is gzus.}\label{johannes-is-gzus.}}

\hypertarget{intended-extensions-to-be-implemented.}{%
\chapter{3. Intended extensions to be
implemented.}\label{intended-extensions-to-be-implemented.}}

\hypertarget{comments}{%
\chapter{4. Comments}\label{comments}}

\hypertarget{references}{%
\chapter*{5. References}\label{references}}
\addcontentsline{toc}{chapter}{5. References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-R6}{}%
Chang, Winston. 2019. \emph{R6: Encapsulated Classes with Reference
Semantics}. \url{https://CRAN.R-project.org/package=R6}.

\leavevmode\hypertarget{ref-Fahrmeir.2013}{}%
Fahrmeir, Ludwig. 2013. \emph{Regression: Models, Methods and
Applications}. New York: Springer.

\leavevmode\hypertarget{ref-Hastie.2009}{}%
Hastie, Trevor, Robert Tibshirani, and J. H. Friedman. 2009. \emph{The
Elements of Statistical Learning: Data Mining, Inference, and Prediction
/ Trevor Hastie, Robert Tibshirani, Jerome Friedman}. 2nd ed. Springer
Series in Statistics. New York: Springer.

\leavevmode\hypertarget{ref-R-asp20model}{}%
Riebl, Hannes. 2020. \emph{Asp20model: An R6 Class for Location-Scale
Regression Models}.

\end{document}