### 2020-05-27 Reports/test_report.Rmd

---
title: ASP20 Boost
subtitle: jai jai
titlerunning: Short form of title (if too long for head)
authorrunning:
  Short form of author list if too long for running head
thanks: |
  Grants or other notes about the article that should go
  on the front page should be placed here. General
  acknowledgments should be placed at the end of the article.
authors:
  - name: Author 1
    address: Department of YYY, University of XXX
    email: abc@def
  - name: Author 2
    address: Department of ZZZ, University of WWW
    email: djf@wef
keywords:
  - key
  - dictionary
  - word
MSC:
  - MSC code 1
  - MSC code 2    
abstract: |
  The text of your abstract.  150 -- 250 words.
bibliography: citavi_27.05.bib
output: rticles::springer_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'dlstats', 'randomForest', 'tidyverse'  
), 'packages.bib')
```