### 2020-05-27 Reports/1_REPORT.Rmd
- Reformated outline - Expanded the introduction.

---
title: "1 Report ASP20 Boost"
subtitle: "First Report"
author:
- "Johannes Strauß"
- "Levin Wiebelt"
- "Sebastian Aristizabal"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: [nothing, nothingness]
abstract: |
  This is the abstract.

  It consists of two paragraphs.
documentclass: report
bibliography: ["citavi_27.05.bib", "packages.bib"]
biblio-style: apalike
geometry: "left=4cm, right=3cm, top=2.5cm, bottom=2.5cm"
lof: FALSE
lot: FALSE
output: 
    pdf_document:
      toc: true
      toc_depth: 2
      latex_engine: xelatex
      keep_tex: yes
      template: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# How to cite and configure the document's appearance 
If you don't have Latex download latex for R with `tinytex::install_tinytex()`.

## Write

To cite:

* Parentheses: Implementation builds upon the asp20 model [@R-asp20model]. 
* Inline: It is implemented as R6 class @R-R6.
* Including pages: For the theoretical foundations of our implementation, we rely mostly on  [@Fahrmeir.2013, 219] and [@Hastie.2009, 358].

In the `report` folder there's two `.bib` files. 
These contain all reference we should need. 
To update the `package.bib` write the name of the package in the following chunk and *actively* run it (It should be installed in your terminal).
The`citavi.bib` file needs to be exported from citavi.

```{r pack tex, eval=FALSE, include=TRUE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'R6', 'knitr', 'rmarkdown', 'asp20model', 'gamboostLSS','mboost', 'tidyverse'  
), 'packages.bib')
```

To cite write a ``@` followed by `firstname.year` as shown in the `.bib` files for reference. 
See further [cite](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)

As a tip hit always enter after each period for better readability in R Studio.
Like this.

Two enters give me a new paragraph.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

## Appearance using the `YAML` Header:

We can regulate the appearance of the document from here. 
There's a lot flexibility but the learning curve is somehow steep.
I'll link useful resources to learn, but, in principle, the document is *ready for writing*.

Here the corresponding references:

*  [PDF output](https://bookdown.org/yihui/rmarkdown/pdf-document.html)
*  [Pandoc Manuals](https://pandoc.org/MANUAL.html#templates)
*  [TOC](https://bookdown.org/yihui/rmarkdown/pdf-document.html#table-of-contents-1)
*  [Cite with Pandoc](https://github.com/jgm/pandoc-citeproc/blob/master/man/pandoc-citeproc.1.md)

Here those inputs that are easy to configuration for you guys to get an idea:

* With `documentclass` has three base templates: `article`, `report` and `book`. 
* `geometry` let's me change the margins. 
* `lof` and `lot` set to true give me a list of figures and tables respectively.

# 1. Introduction:

* The developement of this package takes place in the framework of the seminar "Advanced Statisical Programming" -> SS20
* We expand the R6 class *"LocationScaleRegression"* belonging to the `asp20model` package concieved specifficaly for this seminar [@R-asp20model]
* Our aim is to develope an multi-faceted implementation of boosting for location and scale including component-wise boosting, use of cross-validation to determine the optimal stopping critera  and useful options for visualization.

# 2. Description of progress

> Hannes lead question: What are the critical aspects of your implementation and your package? Stability? Performance? User-friendliness? How are you going to deal with these potential issues?

> 2. What design designs did you make with respect to your implementation? How are you going to extend the R6 class? What additional functions are you going to provide?

I propose a timeline where the implementation process is described step by step as a list e.g:

1. Primitive implementation of boosting
2. Simple boosting for location implemented
3. Adressed the estimation for the variance of gamma. 
4. Johannes finished everything #lol. 27.05.

## Johannes is *gzus*.

# 3. Description of the problem

> Hannes lead question: Which open questions do you have about the statistical model and methodology?

# 4. Extensions to be implemented.

> Hannes' Lead question: What functionality are you planning to (or did you already) implement? What did you decide to leave out?

# 5. Summary?
# 6. References