### 2020-07-24 Reports/report_2/2_REPORT.Rmd
Levin's trasnscript. Spell Check. Formatting. First knitt.

---
title: "ASP20 Boost"
author:
- Johannes Strauß
- Levin Wiebelt
- Sebastian Aristizabal
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    keep_tex: no
    latex_engine: xelatex
    template: null
    toc: yes
    toc_depth: 2
documentclass: report
geometry: left=4cm, right=3cm, top=2.5cm, bottom=2.5cm
lof: FALSE
lot: FALSE
bibliography:
- citavi.bib
- packages.bib
subtitle: Second Report
biblio-style: apalike
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r pack tex, include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'R6', 'knitr', 'rmarkdown', 'asp20model', 'gamboostLSS','mboost', 'tidyverse'  
), 'packages.bib')
```

# 1. Introduction

Following up on the first report from June 5th 2020, this document describes the further progress of developing the R-package asp20boost.
In part 2, we describe the changes in design and new features since completion of the first report, these include amongst others automated unit tests and the improvement of matrix calculations via Cholesky-decomposition.
To assess the performance of the gradient-boost algorithm, we conducted several simulation studies.
The results are described in part 3. Finally, in part 4, we outline which challenges remain.

# 2. Code Updates

## Cholesky Decomposition

Following the issue raised by Hannes Riebl in our Gitlab-Repo, we implemented a Cholesky decomposition in the matrix calculations of the `LocationScaleRegressionBoost` class.
As described in the first report the idea of boosting consists in the iterative estimation of
* Residuals in the case of location-estimation
* The gradient (wrt. scale parameters) of the loss-function in the case of scale estimation.

In both cases this is done by the least squares procedure [insert formula: (X’X)^(-1) * X‘ * residuals_vector]. 
If repeatedly executed, this calculation may be improved by decomposing the symmetric positive definite matrix X’X via Cholesky's procedure.
The inverse then is calculated rapidly by the R-command `chol2inv()`.

## Automated Unit Tests

To automate the testing process of our package we began with the implementation of unit tests.
Up to now, we automated the check for correct model class creation.
In order to test the boosting algorithm with simulated data, external R-scripts seem more adequate.

## API Design

Componentwise boosting demands selection of the component, which is to be updated.
Both this selection and the ensuing calculation of the update are done by the fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.
We decided to move these calculations from active fields to public fields of the `LocationScaleRegressionBoost` class.
This is motivated by the fact that a substantial amount of calculations happen in these fields, while active fields should be of low computational cost.
Hence public fields are more accurate.

# 3. Simulation Studies

In order to assess the functionality of the `gradient_boost` function, we conducted several simulation studies which may be categorized into two sets.
The first set is implemented in the file *simulation_plot_4_params.R*.
It sets up a simple location-scale model with two parameters beta for location, as well as two parameters gamma for scale.
The algorithm does not yet work properly for all possible parameter combinations of beta and gamma.
The second set is implemented in the file *simulation_variable_selection.R*.
It aims to check if variable selection works correctly for multidimensional (dim > 2) beta or gamma.
Results also depend on the concrete parameter configuration.

## Two-dimensional Case

Setting up the model identically to the example in the *asp20model-Repo* lead to algorithm convergence towards the true values beta = (0, 1)‘ and gamma = (-3, 2)‘.

Let the true model be characterized by the following parameter configuration: beta = (1, 1)‘ and gamma = (1, 1)‘. 
After three iterations, the algorithm yields estimates of beta_hat = (0.01, 0.05)‘ and gamma_hat = (12.30, 0.00)‘.
The algorithm detects no further change in gradients and stops prematurely.
By lowering the `abstol` parameter functionality may be restored.
After 3412 iterations estimates are beta_hat = (0.80, 1.44)‘ and gamma_hat = (0.96, 1.12)‘.
These values are not yet good estimates but reasonably close to conclude basic functionality of the boosting algorithm.

In the case of parameter configuration : beta = (2, 2)‘ and gamma = (2, 2)‘ early stopping remains as a problem. 
Functionality can not be achieved by lowering the tolerance threshold `abstol` of the break condition.
The break condition is a comparison of gradients of successive iterations.

## Multidimensional Case

In the simulation studies for multidimensional models, a total of six parameters for beta as well as six for gamma may be specified.

### Simple Case

A simple case consists in setting intercept and one slope for beta and gamma to one while the remaining 2x4 parameters are zero: beta = (1, 1, 0, 0, 0, 0)‘ and gamma = (1, 1, 0, 0, 0, 0)‘.
The algorithm should depict the 2x2 parameters of interest and estimate them close to one while the others should remain zero.
The resulting estimates displayed below indicate proper estimates for the scale parameter vector `gamma_hat` but not for the location estimates `beta_hat`.
```
model$beta
[1]  0.15122097  1.72916260  0.61363927 -0.14620695  0.02756381  0.28560006
> model$gamma
[1]  0.95706770  1.15196564 -0.02756374  0.00000000 -0.06839582  0.00000000
```
```
model$beta
[1]  0.15122097  1.72916260  0.61363927 -0.14620695  0.02756381  0.28560006
model$gamma
[1]  0.95706770  1.15196564 -0.02756374  0.00000000 -0.06839582  0.00000000
```
### Example Case

Setting up the example model from the asp20model repo, the algorithm, in turn, finds correct estimates – however being somewhat inaccurate for gamma_1_hat:
```
model$beta
[1] -0.0148539032  1.0130172152  0.0108580428 -0.0026268607  0.0001592711
[6]  0.0049089286
model$gamma
[1] -3.103392229  1.258857920  0.017474511 -0.005902031 -0.092450456
[6]  0.000000000
```
### 1-0-Shifted Case

Another model setup beta = (1,0,1,0,1,0)‘ and gamma = (0,1,0,1,0,1)‘ yields again correct estimates for gamma, but not for beta:

```
model$beta
[1]  0.69898406  0.53163733  1.64776999 -0.92116479  1.10914562 -0.02877867
model$gamma
[1]  0.01368480  1.17157405 -0.03057068  0.94892465 -0.03256148  0.89579817
```

## Conclusions of the simulation studies

From these findings we conclude that the evolution of gradients during iterations needs to be checked.
We need to find out why gradients converge without yielding sufficiently adequate parameter estimates.
At the moment component-wise boosting is hardwired into our code.
A useful implementation would be to allow for non-componentwise boosting again.
Being able to switch between overall and componentwise boosting may facilitate locating weak points in our code.

# 4. Further Tasks

## Ensure functionality of gradient-boost algorithm

As depicted in part 3, functionality of our algorithm needs to be ensured independent of the concrete data-generating process.

## Choice of initial parameter values

The performance of the boosting algorithm may be enhanced if initial estimates are chosen sensibly.
For the location parameters beta this may be the OLS-estimate or the mean response. For the scale parameters gamma this may be the (homoskedastic) sample variance.
Another possibility is to ask other student groups for straight-forward and fast numerical procedures to determine good starting values.
However, this may be costly in terms of computation time and hence slow down overall performance of the algorithm.
Trade-offs need to be considered.