### 2020-09-15 simulation_studies/sim_study.Rmd
update simulation studies

---
title: "R Notebook"
output: html_notebook
---

```{r}
# run all code chunks in this document
# do not display output of this codechunk
# DO display code
rm(list=ls())
options(scipen=999)
source("helper_plot_model.R")
source("helper_init_model.R")
source("helper_calc_mse.R")
source("helper_init_large_model.R")
library(asp20boost)
library(gamboostLSS)
library(microbenchmark)
```

## Benchmark gamboostLSS

In order to assess performance of our algorithm further we compare it to the existing package *gamboostLSS.* 
Therefore we use configurations 1 and 4 the two-dimensional cases.
In the former case our algorithm performed well, in the latter rather bad.

### gamboostLSS- package

In the *gamboostLSS* package boosting a location-scale regression model works as follows.
First we need to define the vector of controls.
As the package contains more functionalities than ours, we have to specify a control variable, that matches stepsize, and mstop to our package to allow for comparison.

```{r}
ctrl <- boost_control(trace = FALSE,
                      mstop = c(mu = 1000, sigma = 1000),
                      nu = c(mu = 0.01, sigma = 0.001)
)
```

### Recap Configuration 1

```{r}
set.seed(1337)
init_1 <- init_model(beta_1, gamma_1, 1000)$model
mod_1 <- init_1$model
response_1 <- init_1$response
covariate_1 <- init_1$covariate
gamboostLSS(formula = response_1 ~ covariate_1,
                      control = ctrl)
```

What we get as result are not yet parameter estimates, but linear predictors.
These must be regressed on the parameters to arrive at the desired estimates.
In this package we do not first create a model and then boost it. Creation and boosting is performed in one call, already yielding estimates. 
These estimates are, however not beta and gamma, but rather eta_mu and eta_gamma. 
So these unit-wise linear predictors need to be regressed on the covariate in order to arrive at the estimates of interest.

```{r}
eta_mu <- fitted(est_1)$mu
eta_sigma <- fitted(est_1)$sigma

est_beta <- lm(eta_mu ~ covariate_1)
est_gamma <- lm(eta_sigma ~ covariate_1)

```

### Recap Configuration 3

```{r}
init <- init_model(beta_3, gamma_3, 1000)
mod_3 <- init$model
response_3 <- init$response
covariate_3 <- init$covariate
plot_model(beta_3, gamma_3, 1000)
```
[...]

```{r}
# estimate with external package --------

mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3,
                      control = ctrl)
eta_mu <- fitted(mod_33)$mu
eta_sigma <- fitted(mod_33)$sigma

lm(eta_mu ~ covariate_3)
lm(eta_sigma ~ covariate_3)

```

[...]
```{r}
gradient_boost(mod_3)
round(mod_3$beta, 3)
round(mod_3$gamma, 3)
```

```{r}
microbenchmark(
  gradient_boost(mod_3)
)
```

```{r}
microbenchmark(
  mod_33 <- gamboostLSS(formula = response_3 ~ covariate_3, control = ctrl),
  eta_mu <- fitted(mod_33)$mu,
  eta_sigma <- fitted(mod_33)$sigma,
  lm(eta_mu ~ covariate_3),
  lm(eta_sigma ~ covariate_3)
)
```

Ensuring comparability is a difficult task, since the two packages are implemented in rather different ways. Running gradient_boost takes about 657000 microseconds (0.66 seconds). In contrasts, running the external algorithm gamboostLSS takes 2.829.493 microseconds (2.8 seconds), which is a multiple of our algorithm. Additionally, beta and gamma parameters are not yielded directly, but need to be determined by regression as well. In contrats, this happens directly in our algorithm.

To be fair we evaluate construction of the model object as well.

```{r}
microbenchmark(
  mod_1 <- LocationScaleRegressionBoost$new(response_3 ~ covariate_3, ~ covariate_3)
)
```

This changes the situation only marginally, as this is a quite fast step.