### 2020-09-15 simulation_studies/sim_study.Rmd
vignette almost finished + PDF. Missing abstract and conclusion

---
title: "Simulation Study"
output: html_notebook
---

In this part we design and perform a simulation study in order to demonstrate functionality of the implemented boosting algorithm. 
For this, we first employ models with two location and scale parameters each, and show basic functionality. 
Second, validity of the variable selection features is demonstrated by using location and scale parameters with up to 50 components.
Lastly, we compare our algorithm with the one implemented in the `gamboostLSS` package. *really?*

```{r setting, cache=TRUE}
rm(list=ls())
options(scipen=999)
source("../simulation_studies/helper_plot_model.R")
source("../simulation_studies/helper_init_model.R")
source("../simulation_studies/helper_calc_mse.R")
source("../simulation_studies/helper_init_large_model.R")
library(asp20boost)
library(gamboostLSS)
library(microbenchmark)
```

## Two-dimesional Case

The goal in this first part is to demonstrate that the algorithm finds correct estimates for sensible choices of $\bm{\beta}$ and $\bm{\gamma}$.
We use the following 4 different parameter configurations to explore location-scale regression model characteristics.

```{r ,include=FALSE, echo=FALSE, cache=TRUE}
# config 1 ----------------------
beta_1 <- c(0,1)
gamma_1 <- c(-3,2)

# config 2 ----------------------
beta_2 <- c(2,2)
gamma_2 <- c(2,2)

# config 3 ----------------------
beta_3 <- c(0,1)
gamma_3 <- c(-1,-0.3)

# config 4 ----------------------
beta_4 <- c(0,15)
gamma_4 <- c(1,-1)
```

```{r, cache=TRUE}
set.seed(1337)
mod_1 <- init_model(beta_1, gamma_1, 1000)$model
gradient_boost(mod_1)
mod_2 <- init_model(beta_2, gamma_2, 1000)$model
gradient_boost(mod_2)
mod_3 <- init_model(beta_3, gamma_3, 1000)$model
gradient_boost(mod_3)
mod_4 <- init_model(beta_4, gamma_4, 1000)$model
gradient_boost(mod_4)
```

```{r, cache=TRUE, fig.height = 8, fig.width = 11}
# display code
# display output
par(mfrow = c(2,2))
plot_model(beta_1, gamma_1, 1000, "Config 1")
plot_model(beta_2, gamma_2, 1000, "Config 2")
plot_model(beta_3, gamma_3, 1000, "Config 3")
plot_model(beta_4, gamma_4, 1000, "Config 4")
```

```{r, include=FALSE, cache=TRUE}
# We need output. IMPORTANT

result_df <- data.frame("Config" = c(1,1,2,2,3,3,4,4),
                        "beta_true" = c(beta_1, beta_2, beta_3, beta_4),
                        "beta est." = c(round(mod_1$beta, 3), round(mod_2$beta, 3), round(mod_3$beta, 3), round(mod_4$beta, 3)),
                        "gamma_true" = c(gamma_1, gamma_2, gamma_3, gamma_4),
                        "gamma est." = c(round(mod_1$gamma, 3), round(mod_2$gamma, 3), round(mod_3$gamma, 3), round(mod_4$gamma, 3))
                        )
result_df
```

### Evaluation Configuration 1

Values: $\bm{\beta} = (0,1)'$ and $\bm{\gamma} = (-3,2)'$

This parameter configuration is equivalent to the base example case of the `asp20model` package. 
As already mentioned, heteroskedasticity is present, as indicated visually by the funnel shape of the scatter plot.
The solid black line represents the true response mean, while the dashed lines represent the true upwards and downwards standard deviations from this mean.
The red arrow indicates the standard deviation at the point $x = 1$ i.e. the upper limit of covariate values, since covariates are designed to be on the unit interval.
For cases with increasing scale, the red arrow hence indicates maximum standard deviation of the response.
The blue arrow depicts the slope of the location parameter vector $\bm{\beta_1}$. 
The ratio of those two quantities is essential for the estimability of the model. 
Here, the blue arrow is dominant indicating that the model may be well estimated. 
Estimates are displayed in the table above. 
We see that both $\bm{\beta}hat$ and $\hat{\bm{\gamma}}$ estimates are close to the true parameter vectors $\bm{\beta}$ and $\bm{\gamma}$.

### Evaluation Configuration 2

Values: $\bm{\beta} = (2,2)'$ and $\bm{\gamma} = (2,2)'$

Configuration 2 is a somewhat naive parameter configuration where all parameters are set to value $2$. 
Estimating this model does not yield valid results.
The `gradient_boost()` function runs into numerical problems.
To understand why this happens, we need to understand the model characteristics resulting from parameter configuration `2-2-2-2`.
Here, the red arrow shows a large size, while the blue arrow is hardly visible. 
This happens because of the huge standard deviations present in response $y$ that concealing the changes in the mean $\bar{y}$. 
In fact, the standard deviation at $x = 1$ is $exp(2+2) = 54.60$.
In contrast, the true response mean values range from $0$ to $2$.
The change in response location (the signal), indicated by the blue arrow, perishes *stirbt?* due to the high amount of noise caused by the high $\bm{\gamma}$ parameters. 
Thus, a signal-noise ratio results that is impossible to estimate. 
Values for parameters $\bm{\gamma}$ need to be chosen sensibly for simulating data. 
Nevertheless, this is no straight-forward task, since the effect of $\bm{\gamma}$ on the response standard deviation is not linear but exponential. 
To check this *check the expoentiality?* in the two-dimensional case, the arrows in the model plot may be of help.

#### Evaluation Configuration 3

Values: $\bm{\beta} = (0,1)'$ and $\bm{\gamma} = (-1,-0.3)'$

Configuration 3, in turn, yields reasonable estimates.
It contains a negative $\bm{\gamma}$ slope leading to decreasing response standard deviations.
This can be observed by the funnel shape of the plot pointing in the opposite direction as before.

It may be of interest to check if estimates lie closer to the true parameters relative to  e.g. configuration 1.
To do this, we measure the gap between true values and estimates by the mean squared error *why not MSE*.
This enables comparison of configurations 1 and 3.

```{r, cache=TRUE}

mse_1 <- calc_MSE(beta_1, gamma_1, mod_1$beta, mod_1$gamma)
mse_3 <- calc_MSE(beta_3, gamma_3, mod_3$beta, mod_3$gamma)
cbind("config 1" = mse_1,
      "config_3" = mse_3)
```

Configuration 1 leads to smaller MSE. 
This indicates a better performance in this  configuration.

#### Evaluation Configuration 4

Values: $\bm{\beta} = (0,15)'$ and $\bm{\gamma} = (1,-1)'$

In configuration 4, we again included both a negative slope for $\bm{\gamma}$  and a very high slope $\bm{\beta}$.
The idea is to ensure a very good signal-noise ratio to guarantee a  good model fit measured by a low MSE.
Contrary to our expectations, the algorithm performs poorly for this parameter configuration.

```{r, cache=TRUE}
mse_4 <- calc_MSE(beta_4, gamma_4, mod_4$beta, mod_4$gamma)
cbind("config 1" = mse_1,
      "config_3" = mse_3,
      "config 4" = mse_4)
```

The bad model fit is due to the fact that the location slope is highly underestimated.

### Conclusions

The observed two-dimensional cases let us conclude that not every arbitrary parameter configuration is estimable by the implemented `gradient_boost()` algorithm. 
A first insight is that signal-noise ratio needs to be sufficiently high in order to prevent that the location effect gets concealed by inflated standard deviations. 
Regarding the finding of configuration 4, we offer several explanations:

  * Further considerations to  simulating data sensibly need to be taken,  which we were not yet able to work out.
  * Our algorithm does not reliably yield robust results if negative $\bm{\gamma}$ slopes are included.
  * Our algorithm runs into problems if the signal is too strong compared to noise.

## Large Model - Variable Selection in the multidimensional case

In this part we demonstrate how the `gradient_boost()` function performs variable selection.
We have a look at three large model configurations.
Configuration 5 present 6-dimensional parameter setting.
In configuration 6 $\bm{\beta}$ and $\bm{\gamma}$ are of length 11, while in configuration 7 they contain 51 components.
The parameter vectors are created randomly. 
Then some components are set to 0. 
This results in the configurations displayed below.
In order to ensure a good signal-noise ratio, to creation of random parameter is restricted, as explained below.

```{r, echo=FALSE, cache=TRUE}
set.seed(1337)
# initialize models --------------------------------------------------------------
init_5 <- init_large_model(beta_range = c(5,10),
                           gamma_range = c(0.5,1),
                           dim_param = 5,
                           int_gamma = -0.2,
                           n = 100)
init_6 <- init_large_model(beta_range = c(3,6),
                           gamma_range = c(0.2,0.8),
                           dim_param = 10,
                           int_gamma = -0.2,
                           n = 100)
init_7 <- init_large_model(beta_range = c(5,10),
                           gamma_range = c(0.01,0.08),
                           dim_param = 50,
                           int_gamma = 0,
                           n = 100)
mod_5 <- init_5$model
mod_6 <- init_6$model
mod_7 <- init_7$model

# estimate models -----------------------------------------------------------------
gradient_boost(mod_5, stepsize = c(0.1,0.001), maxit = 1000, componentwise = TRUE)
gradient_boost(mod_6, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)
gradient_boost(mod_7, stepsize = c(0.1,0.01), maxit = 1000, componentwise = TRUE)

```

```{r}
#dont show code
#show output
# Configuration 5
cbind(beta_true = init_5$beta, 
      beta_est = round(mod_5$beta, 3),
      gamma_true = init_5$gamma,
      gamma_est = round(mod_5$gamma, 3)
)
```

```{r echo=FALSE, cache=TRUE}
#dont show code
#show output
# Configuration 6
cbind(beta_true = init_6$beta, 
      beta_est = round(mod_6$beta, 3),
      gamma_true = init_6$gamma,
      gamma_est = round(mod_6$gamma, 3)
)
```

### Configuration 5

As mentioned above there are restrictions on parameter creation since not every possible parameter configuration leads to estimable models, as we have seen in the two-dimensional case.
In the multidimensional case, the following  problem remains: if $\bm{\gamma}$ parameters add up to high values, the resulting simulated response has huge standard deviation. 
This leads to the concealing of the location effects.
Checking for well balanced signal-noise ratio remains to be done in the multidimensional case as well.
Furthermore, plotting is hardly an option, as dimensions increase.
Therefore we check using another method i.e. we look at the distributions of true response means and true response standard deviations.

```{r, cache=TRUE}
par(mfrow=c(2,1))
hist(init_5$fitted_response, breaks = 25, main = "eta_mu")
hist(init_5$fitted_sd, breaks = 25, main = "exp_eta_sigma")
```

Location predictors take values between 8 and 22 with a maximum of 15. 
In contrast, the highest response- standard deviation is 4, while most values are at about 2.5 and lower. 
The resulting response variation should be both small enough to not conceal the location effects and big enough to be estimated.

In configuration 5, the algorithm correctly depicts which effects exists on the location response *location response oder response location?* i.e. it finds the non-zero elements of $\bm{\beta}$.
For the scale estimate $\bm{\gamma}_hat$ too many parameters are chosen since two zero elements of the true parameter $\bm{\gamma}$ are assigned non-zero values by the estimate $\hat{\bm{\gamma}}$.
The estimated effect sizes differ considerably higher from the true values relative the two-dimensional case already presented.  This results in a higher MSE.

```{r, cache=TRUE}
mse_5 <- calc_MSE(init_5$beta, init_5$gamma, mod_5$beta, mod_5$gamma)
cbind("config 1" = mse_1,
      "config_3" = mse_3,
      "config 4" = mse_4,
      "config_5" = mse_5)
```

### Configuration 6

In this case, we face a 10-dimensional configuration.
We get the following distributions for the mean and the standard deviation respectively.

```{r, cache = TRUE}
par(mfrow=c(2,1))
hist(init_6$fitted_response, breaks = 25, main = "eta_mu")
hist(init_6$fitted_sd, breaks = 25, main = "exp_eta_sigma")
```

For $\bm{\beta}$ the algorithm chose 1 *what?* too few.
For $\bm{\gamma}$ entries 1-3 are chosen correctly.
Then the algorithm yields errors.
Entry 4 is chosen, while in fact being zero.
Entry 5 is not chosen even when an effect in the true model exists.

### Configuration 7

When scaling the parameter dimension up to 51 variables, the algorithm runs into a critical problem: all slope estimates are zero.

```{r, cache = TRUE}
round(mod_7$beta, 3)
round(mod_7$gamma, 3)
```

Histograms indicate a well-balanced signal-noise ratio. 
The problem's cause lies thus somewhere else.

```{r, cache = TRUE}
par(mfrow=c(2,1))
hist(init_7$fitted_response, breaks = 25, main = "eta_mu")
hist(init_7$fitted_sd, breaks = 25, main = "exp_eta_sigma")
```

One explanation for this problem is that the algorithm fits only intercepts in the first step resulting in very high values.
The remaining iterations are used to lower intercepts *Um den intercepts to reducing? oder was meints du?*. 
This may be checked by setting `verbose = TRUE`.
For this reason, no other components of the parameter estimates are updated.
Possibly, a sensible choice of initial parameter values may address this problem.